Preliminary Insights into Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan

### 2025-06-01 analysis/paper/paper.qmd
added a reference

The faunal assemblage from the sampling area of Kiwulan consists of `r total_NISP` bone fragments, of which `r family_ident` were identified to at least the family level (@tbl-counts). Among those bones identified to the family or more specific level, Cervidae is the majority (`r cer_NISP_per`%), followed by Suidae (`r suid_NISP_per`%) and Bovinae (`r buff_NISP_per`%). To further compare general patterns of animal consumption based on economic values, animals from the same family and of similar body size were grouped together. Therefore, the deer category includes sika deer, sambar deer, and large deer, while water buffalo and yellow cattle were grouped into the bovine category. Similarly, domestic pigs (CL4: n = 2, CL6: n = 1) were integrated into the suid category. Rats (*Rattus sp.*) were excluded from further analyses as they are usually considered commensal animals, and the relatively intact skeletons recovered from a single archaeological context show no direct of being used for dietary purposes. In general, the faunal remains show that frequencies increased from CL6, peaked at CL4, and then declined dramatically after CL4 (@tbl-counts, @fig-NISP-proportion: A). A similar trend was observed in the weight of unidentified small bone fragments (@fig-NISP-proportion: B).

>>>>

The faunal assemblage from the sampling area of Kiwulan consists of `r total_NISP` bone fragments, of which `r family_ident` were identified to at least the family level (@tbl-counts). Among those bones identified to the family or more specific level, Cervidae is the majority (`r cer_NISP_per`%), followed by Suidae (`r suid_NISP_per`%) and Bovinae (`r buff_NISP_per`%). To further compare general patterns of animal consumption based on their economic values and meat return rates, animals of similar body size and from the same family were grouped together. For example, the "deer" category includes only medium and large deer species, while the smaller deer species, muntjac, which implies a lower meat yield, was excluded and treated as a separate category. Similarly, water buffalo and yellow cattle, which are comparable in size, were grouped into the "bovine" category, and domestic pigs (CL4: n = 2; CL6: n = 1) were included in the "suid" category. Rats (*Rattus sp.*) were excluded from further analyses as they are usually considered commensal animals, and the relatively intact skeletons recovered from a single archaeological context show no direct of being used for dietary purposes. In general, the faunal remains show that frequencies increased from CL6, peaked at CL4, and then declined dramatically after CL4 (@tbl-counts, @fig-NISP-proportion: A). A similar trend was observed in the weight of unidentified small bone fragments (@fig-NISP-proportion: B).
